---
title: "Assignment"
author: "Abhijeet Sharma"
date: "07/12/2020"
output: html_document
---
```{r}
library(dplyr)
library(ggplot2)
doc <- read.csv("activity.csv")
act.complete <- na.omit(doc)
```
###Total Number of Steps Taken Per Day
```{r}
act.day <- group_by(act.complete, date)
act.day <- summarize(act.day, steps=sum(steps))
summary(act.day)
head(act.day)
```
####Histogram
```{r}
hist(act.day$steps, xlab = "Total daily Steps",main="Histogram of Total Steps by day", breaks = 20)
```
####Mean and Median of the total number of steps taken per day
```{r}
mean(act.day$steps)
median(act.day$steps)
```
####Average Daily Activity Pattern
```{r}
act.int <- group_by(act.complete, interval)
act.int <- summarize(act.int, steps=mean(steps))
ggplot(act.int, aes(interval, steps)) + geom_line()
act.int[act.int$steps==max(act.int$steps),]
```
###Imputing Missing Values
```{r}
nrow(doc)-nrow(act.complete)
names(act.int)[2] <- "mean.steps"
act.impute <- merge(doc, act.int)
```
####New dataset that is equal to the original dataset but with the missing data filled in.
```{r}
act.impute$steps[is.na(act.impute$steps)] <- act.impute$mean.steps[is.na(act.impute$steps)]
```
####Histogram of the total number of steps taken each day
```{r}
act.day.imp <- group_by(act.impute, date)
act.day.imp <- summarize(act.day.imp, steps=sum(steps))

qplot(steps, data=act.day.imp)
```
###Differences in activity patterns between weekdays and weekends
####To create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
act.impute$dayofweek <- weekdays(as.Date(act.impute$date))
act.impute$weekend <-as.factor(act.impute$dayofweek=="Saturday"|act.impute$dayofweek=="Sunday")
levels(act.impute$weekend) <- c("Weekday", "Weekend")
```
####Panel plot containing a time series plot 
First I create separate data frames for weekends and weekdays:

```{r}
act.weekday <- act.impute[act.impute$weekend=="Weekday",]
act.weekend <- act.impute[act.impute$weekend=="Weekend",]
```
Then for each one, I find the mean number of steps across days for each 5 minute interval:

```{r}
act.int.weekday <- group_by(act.weekday, interval)
act.int.weekday <- summarize(act.int.weekday, steps=mean(steps))
act.int.weekday$weekend <- "Weekday"
act.int.weekend <- group_by(act.weekend, interval)
act.int.weekend <- summarize(act.int.weekend, steps=mean(steps))
act.int.weekend$weekend <- "Weekend"
```
I append the two data frames together, and I make the two time series plots:
```{r}
act.int <- rbind(act.int.weekday, act.int.weekend)
act.int$weekend <- as.factor(act.int$weekend)
ggplot(act.int, aes(interval, steps)) + geom_line() + facet_grid(weekend ~ .)
```